---
title: "Appendix"
output: html_document
date: "2023-05-08"
---

# Data Imputation 

```{r data_imputation, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#import dataset 
data <- read.csv(here::here("data/heart_disease.csv"))
```

```{r}
library(mice)
library(VIM)
#Plotting missing values 
aggr(data, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing values","Pattern"))
```
```{r}
#Proportion of total missing values 
sum(is.na(data))/nrow(data)
```

:::{style="text-align: justify"}

As we can see missing values represent a high proportion of our data set,almost 15%,most of them come from continuous variables such as "glucose","BPMeds","totChol","cigsPerDay" and "BMI".Therefore we have decided to impute these values using the most appropriate method fo continuous variables : Predictive Mean Matching.The latter consists in using the observed values from the closest matching cases to impute missing values taking into account the correlation between variables 

:::

```{r}
#Method for continuous data
imputed_data <- mice(data,m=19,maxit=50,method = "pmm",seed=500)
summary(imputed_data)
#Check immputed data 
#Given that we have chose m=19 , for every instance where there is missing value we have 19 solutions
imputed_data$imp$glucose
imputed_data$imp$BPMeds
imputed_data$imp$totChol
imputed_data$imp$cigsPerDay
imputed_data$imp$BMI
#Get complete data 
complete_data <- complete(imputed_data,3)
#Optimization for the dataset
head(data,n=15)
#percentage of error : how the predicted data is deviating from the original data 
a=(abs(data[,c("glucose", "BPMeds", "totChol","cigsPerDay","BMI")]-complete_data[,c("glucose", "BPMeds", "totChol","cigsPerDay","BMI")])) #see the difference except the categorical variable education
#Save file and load complete datase
write.csv(x=complete_data,file ="complete_data_MICE.csv")
data <- read.csv("complete_data_MICE.csv")
```

```{r}
#removing missing values from "education" variable 
data <- na.omit(data)
```
