
# **3. Data Modification**

```{r data_modification, echo = FALSE, message = FALSE, warning= FALSE}
#Load imputed data 
data_new <- read.csv(here::here("complete_data_MICE.csv"))
#data_new <- read_csv("Documents/GitHub/ML_Heart-Disease/complete_data_MICE.csv")
```

:::{style="text-align: justify"}
Variables `Gender`,  `prevalentStroke`, and `Heart_stroke` are transformed into numeric dummy variables.

Variable `education` are transformed into Liker scale.

Variables `cigsPerDay`, `totChol`, `sysBP`, `diaBP`, `BMI`, `heartRate`, and `glucose`, `age` are normalized to one scale.

The variable `prevalent_stroke` does not provide useful information to differentiate between different instances in the dataset. Since it has the same value for all instances, it does not contribute to any variation in the data. Therefore, we removed it.
:::

```{r}
# Convert gender to numeric dummy variable

data_new <- data_new %>%
  mutate(gender = ifelse(Gender == "Female", 1, 0),
         heart_stroke = ifelse(Heart_.stroke == "yes", 1, 0),
         prevalent_stroke = ifelse(prevalentStroke == "Yes", 1, 0))

dummys <- data_new %>% select(gender, heart_stroke, prevalent_stroke, BPMeds)


# scale numeric variables
num <- data_new %>% select(cigsPerDay, totChol, sysBP, diaBP, BMI, heartRate, glucose, age) %>% 
  mutate_at(vars(cigsPerDay, totChol, sysBP, diaBP, BMI, heartRate, glucose, age), as.numeric) %>% scale() %>% as.data.frame()


 #Create a liker scale for the education column
liker <- data_new %>%
  mutate(education = ifelse(education  == "uneducated", 1,
                            ifelse(education == "primaryschool", 2,
                                   ifelse(education == "graduate", 3,
                                          ifelse(education == "postgraduate", 4, 0)))))

                          
education <- liker$education

data_transformed_n <- data.frame(dummys, num, education)
```

```{r,echo=FALSE}
plot_histogram <- function(data_transformed_n) {
  
# Plot the histograms
  par(mfrow = c(2, ncol(data_transformed_n)/2)) # set up the grid for plots
  for (i in 1:ncol(data_transformed_n)) {
    hist(data_transformed_n[, i], main = colnames(data_transformed_n)[i], xlab = "", col = "lightblue", border = "white")
  }
}
```

```{r}
#Variables distribution after modification
plot_histogram(data_transformed_n)
```

```{r}
data_transformed <- data_transformed_n %>% 
  select(-prevalent_stroke)
```



